---
title: "Chapter 2 Exercises"
output:
  pdf_document: default
  html_notebook: default
---
```{r}
library(sf)
library(spData)
library(sp)
library(raster)
```
## Question 1
Use summary() on the geometry column of the world data object. What does the output tell us about:
-Its geometry type?
-The number of countries? 
-Its coordinate reference system (CRS)? 
```{r}
summary(world["geom"])
```
Geometry Type: MULTIPOLYGON
Number of Countries: 177 (can be seen from num of polygons)
CRS:  WGS84 (EPSG 4326)

## Question 2 
-What does the cex argument do (see ?plot)?
-Why was cex set to the sqrt(world$pop) / 10000?
-Bonus: experiment with different ways to visualize the global population.

This code was taken from the book but copied into this file for convenience 

```{r}
plot(world["continent"], reset = FALSE)
cex = sqrt(world$pop) / 10000
world_cents = st_centroid(world, of_largest = TRUE)
plot(st_geometry(world_cents), add = TRUE, cex = cex)
```
Differences: This map is a mercator projection while the one in the book isn't, the circles in the book's map are thicker
cex is the symbol size or the size of each circle
cex is set to sqrt(world$pop) / 10000 so the size of the circle is proportionate to the size of each country


## Question 3

```{r}
# take the row associated with nigeria
nigeria = world[world$name_long == "Nigeria", ]
# plot the shape (geometry), set the bounding box, color, and line width
plot(st_geometry(nigeria), expandBB = c(1, 1, 0.5, 1), col = "pink", lwd = 4)
# get all the countries in afria
africa = world[world$continent == "Africa",]
# add all the other countries to the plot
plot(africa[0], add = TRUE)
# who determines what countries are part of of Africa and which borders are used?
# Why is this the default projection?
# challenge-- used text function by specifying loation and the text I wanted displayed
# determined x and y values through trial and error 
text(x = 7, y = 8, labels = "Nigeria")

```

## Question 4
```{r}
# create a blank raster
my_raster = raster(nrows = 10, ncols = 10)
values = vector(mode = "numeric", length = 100)
 for (i in 1:100){
    values[i] = sample(1:10, 1);
 }


# I thought there was a better way to generate random numbers 100 times than a loop so I tried this but it didn't seem to work so I went back to the loop
# sapply(values, sample, 1:10, 1)

# take the values I generated and set raster label values at them 
my_raster = setValues(my_raster, values)

# This line was from when I was trying to make sure my numbers were generating properly
#getValues(my_raster)

# plot raster by specifying my raster, which colors i want, transparency, and the max pixels in the raster
 plot(my_raster, col = hcl.colors(10, palette = "viridis", alpha = NULL), maxpixels = 100)
```

## Question 5

```{r}
# read in the file
# where did this layer come from? who made it? for what purpose?
file = raster(system.file("raster/nlcd2011.tif", package = "spDataLarge"))
#view info
file
```
You can get the class, dimensions, resolution, extent, information such as projection, szone, units, the source, names, and the min max values
