---
title: "Chapter 2 Exercises"
output: html_notebook
---
```{r}
library(sf)
library(spData)
library(sp)
library(raster)
```
## Question 1
Use summary() on the geometry column of the world data object. What does the output tell us about:
-Its geometry type?
-The number of countries? 
-Its coordinate reference system (CRS)? 

```{r}
summary(world["geom"])
```
Geometry Type: MULTIPOLYGON
Number of Countries: 177 (can be seen from num of polygons)
CRS:  WGS84 (EPSG 4326)

## Question 2 
-What does the cex argument do (see ?plot)?
-Why was cex set to the sqrt(world$pop) / 10000?
-Bonus: experiment with different ways to visualize the global population.

```{r}
plot(world["continent"], reset = FALSE)
cex = sqrt(world$pop) / 10000
world_cents = st_centroid(world, of_largest = TRUE)
plot(st_geometry(world_cents), add = TRUE, cex = cex)
```
Differences: This map is a mercator projection while the one in the book isn't, the circles in the book's map are thicker
cex is the symbol size or the size of each circle
cex is set to sqrt(world$pop) / 10000 so the size of the circle is proportionate to the size of each country

## Question 3

```{r}
nigeria = world[world$name_long == "Nigeria", ]
plot(st_geometry(nigeria), expandBB = c(3, 0.2, 0.1, 1), col = "pink", lwd = 4)
africa = world[world$continent == "Africa",]
plot(africa[0], add = TRUE)


```

## Question 4
```{r}
my_raster = raster(nrows = 10, ncols = 10)
values = vector(mode = "numeric", length = 100)
 for (i in 1:100){
    values[i] = sample(1:10, 1);
 }

# I thought there was a better way to generate random numbers 100 times than a loop so I tried this but it didn't seem to work so I went back to the loop
# sapply(values, sample, 1:10, 1)
my_raster = setValues(my_raster, values)
# This line was from when I was trying to make sure my numbers were generating properly
#getValues(my_raster)

 plot(my_raster, col = hcl.colors(10, palette = "viridis", alpha = NULL, rev = FALSE, fixup = TRUE)
, maxpixels = 100)
```

## Question 5

```{r}
nlcd = raster(system.file("raster/nlcd2011.tif", package = "spDataLarge"))
nlcd
```
You can get the class, dimensions, resolution, extent, information such as projection, szone, units, the source, names, and the min max values

