---
title: "Chapter 3"
output: html_notebook
---
0. Load libraries
```{r}
library(spData)
library(dplyr)
library(RQGIS)
data(us_states)
data(us_states_df)
```

1. Create a new object called us_states_name that contains only the NAME column from the us_states object. What is the class of the new object and what makes it geographic?

```{r}
us_states_name= us_states[, 2]
class(us_states_name)
```

2. Select columns from the us_states object which contain population data. Obtain the same result using a different command (bonus: try to find three ways of obtaining the same result). Hint: try to use helper functions, such as contains or starts_with from dplyr (see ?contains).

```{r}
population1 = us_states[ , 5:6]

population2 = data.frame(us_states$total_pop_10, us_states$total_pop_15)
```

3. Find all states with the following characteristics (bonus find and plot them):
Belong to the Midwest region.
Belong to the West region, have an area below 250,000 km2 and in 2015 a population greater than 5,000,000 residents (hint: you may need to use the function units::set_units() or as.numeric()).
Belong to the South region, had an area larger than 150,000 km2 or a total population in 2015 larger than 7,000,000 residents.

```{r}
found_states = us_states[us_states$REGION == "Midwest" | (us_states$REGION == "West" & us_states$AREA  < units::set_units(250000, km^2) |us_states$total_pop_15 > 5000000) | (us_states$REGION == "South" & (us_states$AREA < units::set_units(150000, km^2)|us_states$total_pop_15 > 7000000)),]

```

4. What was the total population in 2015 in the us_states dataset? What was the minimum and maximum total population in 2015?
```{r}
sum(us_states$total_pop_15)
min(us_states$total_pop_15)
max(us_states$total_pop_15)
```

5. How many states are there in each region?
```{r}
nrow(us_states[us_states$REGION == "Midwest", ])
nrow(us_states[us_states$REGION == "West", ])
nrow(us_states[us_states$REGION == "Norteast", ])
nrow(us_states[us_states$REGION == "South", ])
```

6. What was the minimum and maximum total population in 2015 in each region? What was the total population in 2015 in each region?
```{r}
sum(us_states[us_states$REGION == "Midwest", ]$total_pop_15)
min(us_states[us_states$REGION == "Midwest", ]$total_pop_15)
max(us_states[us_states$REGION == "Midwest", ]$total_pop_15)

sum(us_states[us_states$REGION == "West", ]$total_pop_15)
min(us_states[us_states$REGION == "West", ]$total_pop_15)
max(us_states[us_states$REGION == "West", ]$total_pop_15)

sum(us_states[us_states$REGION == "South", ]$total_pop_15)
min(us_states[us_states$REGION == "South", ]$total_pop_15)
max(us_states[us_states$REGION == "South", ]$total_pop_15)

sum(us_states[us_states$REGION == "Norteast", ]$total_pop_15)
min(us_states[us_states$REGION == "Norteast", ]$total_pop_15)
max(us_states[us_states$REGION == "Norteast", ]$total_pop_15)
```

7. Add variables from us_states_df to us_states, and create a new object called us_states_stats. What function did you use and why? Which variable is the key in both datasets? What is the class of the new object?
```{r}
us_states_stats = left_join(us_states, us_states_df, by = c("NAME" = "state"))
class(us_states_stats)
```

8. us_states_df has two more rows than us_states. How can you find them? (hint: try to use the dplyr::anti_join() function)
```{r}
anti_join(us_states_df, us_states, by= c("state" = "NAME"))
```

9. What was the population density in 2015 in each state? What was the population density in 2010 in each state?
```{r}
pop_density_15 = us_states$total_pop_15/as.numeric(us_states$AREA)
pop_density_10 = us_states$total_pop_10/as.numeric(us_states$AREA)

pop_density_15
pop_density_10
```

10. How much has population density changed between 2010 and 2015 in each state? Calculate the change in percentages and map them.
```{r}
percent_change = ((pop_density_15 - pop_density_10) / pop_density_10)*100 
percent_change
# don't forget to map 
us_states$percent_change= percent_change
plot(us_states["percent_change"], expandBB = c(.01, .01, 0.01, .01),lwd =1)

```

11. Change the columns’ names in us_states to lowercase. (Hint: helper functions - tolower() and colnames() may help.)
```{r}
names <- colnames(us_states)
colnames(us_states) = tolower(names)
```

12. Using us_states and us_states_df create a new object called us_states_sel. The new object should have only two variables - median_income_15 and geometry. Change the name of the median_income_15 column to Income.
```{r}
us_states_temp = left_join(us_states, us_states_df, by = c("name" = "state"))
us_states_sel = select(us_states_temp, "median_income_15", "geometry")
colnames(us_states_sel)[1] <- "Income"

```

13.Calculate the change in median income between 2010 and 2015 for each state. Bonus: What was the minimum, average and maximum median income in 2015 for each region? What is the region with the largest increase of the median income?
```{r}
income_percent_change = (us_states_df$median_income_15 - us_states_df$median_income_10)/ (us_states_df$median_income_10)*100 
income_percent_change
```

14. Create a raster from scratch with nine rows and columns and a resolution of 0.5 decimal degrees (WGS84). Fill it with random numbers. Extract the values of the four corner cells.

```{r}
random = raster(nrows = 9, ncol = 9, vals = runif(81, 0.0, 1.0), res= c(.5, .5), xmn = 0, ymn = 0, xmx = 4.5, ymx = 4.5)
res(random)
plot(random, col = hcl.colors(10, palette = "viridis", alpha = NULL), maxpixels = 81)
row_1 = getValues(random, row=1)
row_9 = getValues(random, row=9)
row_1[1]
row_1[9]
row_9[1]
row_9[9]
```

15. What is the most common class of our example raster grain (hint: modal())?
```{r}
# this code is from the book to build grain
elev = raster(nrows = 6, ncols = 6, res = 0.5,
              xmn = -1.5, xmx = 1.5, ymn = -1.5, ymx = 1.5, vals = 1:36)
grain_order = c("clay", "silt", "sand")
grain_char = sample(grain_order, 36, replace = TRUE)
grain_fact = factor(grain_char, levels = grain_order)
grain = raster(nrows = 6, ncols = 6, res = 0.5, 
               xmn = -1.5, xmx = 1.5, ymn = -1.5, ymx = 1.5,
               vals = grain_fact)

# apply modal to every factor
cellStats(grain, modal) %>%
  factorValues(grain, .)
factorValues(grain, modal(values(grain)))
```

16. Plot the histogram and the boxplot of the data(dem, package = "RQGIS") raster.
```{r}
hist(dem)
boxplot(dem)
```

