---
title: "Chapter 3"
author: "Charlie Nee"
output:
  html_notebook: default
  pdf_document: default
---
### Reading Response
Monmonier's _How to Lie With Maps_ discusses both the intentional and unintentional ways that maps can be misleading. Monmonier looks at ways scale, projections, and symbols can be misleading. In chapter 3, Monmonier looks at necessary generalizations and how different generalizations can be misleading.Monmonier also gives recommendations to help maps be less misleading. For example, using symbols of different sizes when representing raw numbers and cholorpleths when representing ratios. I also enjoyed the discussion of the DC subway map and appreciated the note that a more "accurate" map may be inappropriate for the situation. \n
I found the discussions such as how different projections emphasize different parts and preserve different features of the world to be interesting, but I hoped that there would be more information on the connection between mapmaking with these different tools and colonialism, as colonialism influnced the way that these potentially misleading tools were developed and these tools were used in order to further colonialism. Finally, I also wish that the state's use of generalized maps got more attention than just a short paragraph at the end, since the state relies so heavily on maps.  \n
0. Load libraries
```{r}
library(spData)
library(dplyr)
library(RQGIS)
data(us_states)
data(us_states_df)
```

1. Create a new object called us_states_name that contains only the NAME column from the us_states object. What is the class of the new object and what makes it geographic?

```{r}
# subset the us states data set by selecting the second column with the names
us_states_name= us_states[, 2]

# check the class
class(us_states_name)
```
The class of the object is a data frame of sf objects. The sticky geometry makes it so that the subset still keeps the geometric data. 


2. Select columns from the us_states object which contain population data. Obtain the same result using a different command (bonus: try to find three ways of obtaining the same result). Hint: try to use helper functions, such as contains or starts_with from dplyr (see ?contains).

```{r}
# subset the us states set directly, select columns 5 and 6
population1 = us_states[ , 5:6]

# create a new data frame with the correct columns
population2 = data.frame(us_states$total_pop_10, us_states$total_pop_15)
```

3. Find all states with the following characteristics (bonus find and plot them):
Belong to the Midwest region.
Belong to the West region, have an area below 250,000 km2 and in 2015 a population greater than 5,000,000 residents (hint: you may need to use the function units::set_units() or as.numeric()).
Belong to the South region, had an area larger than 150,000 km2 or a total population in 2015 larger than 7,000,000 residents.

```{r}
# subset with a vector created with a long sequence of logical operators
found_states = us_states[us_states$REGION == "Midwest" | (us_states$REGION == "West" & us_states$AREA  < units::set_units(250000, km^2) |us_states$total_pop_15 > 5000000) | (us_states$REGION == "South" & (us_states$AREA < units::set_units(150000, km^2)|us_states$total_pop_15 > 7000000)),]

```

4. What was the total population in 2015 in the us_states dataset? What was the minimum and maximum total population in 2015?
```{r}
# adds all of the numbers in the column
sum(us_states$total_pop_15)
# takes the minimum of all the numbers in the column
min(us_states$total_pop_15)
# takes the maximum of all the numbers in the column
max(us_states$total_pop_15)
```

5. How many states are there in each region?
```{r}
# nrow counts the number of rows in a dataframe. Passed in a subsetted by region version of us_states
nrow(us_states[us_states$REGION == "Midwest", ])
nrow(us_states[us_states$REGION == "West", ])
nrow(us_states[us_states$REGION == "Norteast", ])
nrow(us_states[us_states$REGION == "South", ])
```

6. What was the minimum and maximum total population in 2015 in each region? What was the total population in 2015 in each region?
```{r}
# subsetted by region and passed that into the functionns used in question 4
# where did this data come from?
sum(us_states[us_states$REGION == "Midwest", ]$total_pop_15)
min(us_states[us_states$REGION == "Midwest", ]$total_pop_15)
max(us_states[us_states$REGION == "Midwest", ]$total_pop_15)

sum(us_states[us_states$REGION == "West", ]$total_pop_15)
min(us_states[us_states$REGION == "West", ]$total_pop_15)
max(us_states[us_states$REGION == "West", ]$total_pop_15)

sum(us_states[us_states$REGION == "South", ]$total_pop_15)
min(us_states[us_states$REGION == "South", ]$total_pop_15)
max(us_states[us_states$REGION == "South", ]$total_pop_15)

sum(us_states[us_states$REGION == "Norteast", ]$total_pop_15)
min(us_states[us_states$REGION == "Norteast", ]$total_pop_15)
max(us_states[us_states$REGION == "Norteast", ]$total_pop_15)
```

7. Add variables from us_states_df to us_states, and create a new object called us_states_stats. What function did you use and why? Which variable is the key in both datasets? What is the class of the new object?
```{r}
# take the missing columns in us_states that are in us_states_df and add them to us_states
us_states_stats = left_join(us_states, us_states_df, by = c("NAME" = "state"))
class(us_states_stats)
```
I used left join because i wanted to preserve all the info in us_states so i kept that on the left and put the us_states_df on the right. The key for us_states was NAME and the key for us_states_df was "state"


8. us_states_df has two more rows than us_states. How can you find them? (hint: try to use the dplyr::anti_join() function)
```{r}
# why is alaska and hawaii left out?
anti_join(us_states_df, us_states, by= c("state" = "NAME"))
```
Anti join gave me the rows that were not in us_states which was Alaska and Hawaii.

9. What was the population density in 2015 in each state? What was the population density in 2010 in each state?
```{r}
# divide the population each year by the area (the area is the same since the size of the state doesnt change)
pop_density_15 = us_states$total_pop_15/as.numeric(us_states$AREA)
pop_density_10 = us_states$total_pop_10/as.numeric(us_states$AREA)

pop_density_15
pop_density_10
```

10. How much has population density changed between 2010 and 2015 in each state? Calculate the change in percentages and map them.
```{r}
# subtract the population densities from each other and divide by the original one. Multiply by 100 to get a percent
percent_change = ((pop_density_15 - pop_density_10) / pop_density_10)*100 
percent_change

# add percent change to data frame
us_states$percent_change= percent_change

# make the map
plot(us_states["percent_change"], expandBB = c(.01, .01, 0.01, .01),lwd = .5,  col = hcl.colors(10, palette = "purples", alpha = NULL), main = "Population Change 2010-2015")

```

11. Change the columns’ names in us_states to lowercase. (Hint: helper functions - tolower() and colnames() may help.)
```{r}
# save all the current column names
names <- colnames(us_states)

# make all the names lowercase and set them as the new names
colnames(us_states) = tolower(names)
```

12. Using us_states and us_states_df create a new object called us_states_sel. The new object should have only two variables - median_income_15 and geometry. Change the name of the median_income_15 column to Income.
```{r}

# combine us_states and us_states df
us_states_temp = left_join(us_states, us_states_df, by = c("name" = "state"))
# choose only the the columns we want
us_states_sel = select(us_states_temp, "median_income_15", "geometry")

# rename the first column 
colnames(us_states_sel)[1] <- "Income"

```

13.Calculate the change in median income between 2010 and 2015 for each state. Bonus: What was the minimum, average and maximum median income in 2015 for each region? What is the region with the largest increase of the median income?
```{r}

# follow same process as question 9
income_percent_change = (us_states_df$median_income_15 - us_states_df$median_income_10)/ (us_states_df$median_income_10)*100 
income_percent_change
```

14. Create a raster from scratch with nine rows and columns and a resolution of 0.5 decimal degrees (WGS84). Fill it with random numbers. Extract the values of the four corner cells.

```{r}
# genearte new raster with random values
random = raster(nrows = 9, ncol = 9, vals = runif(81, 0.0, 1.0), res= c(.5, .5), xmn = 0, ymn = 0, xmx = 4.5, ymx = 4.5)
# plot the raster
plot(random, col = hcl.colors(10, palette = "viridis", alpha = NULL), maxpixels = 81)
# get only the first and last row
row_1 = getValues(random, row=1)
row_9 = getValues(random, row=9)

# print the first and last value in each of the rows we got 
row_1[1]
row_1[9]
row_9[1]
row_9[9]
```

15. What is the most common class of our example raster grain (hint: modal())?
```{r}
# this code is from the book to build grain
elev = raster(nrows = 6, ncols = 6, res = 0.5,
              xmn = -1.5, xmx = 1.5, ymn = -1.5, ymx = 1.5, vals = 1:36)
grain_order = c("clay", "silt", "sand")
grain_char = sample(grain_order, 36, replace = TRUE)
grain_fact = factor(grain_char, levels = grain_order)
grain = raster(nrows = 6, ncols = 6, res = 0.5, 
               xmn = -1.5, xmx = 1.5, ymn = -1.5, ymx = 1.5,
               vals = grain_fact)

# Look at how many of each value there is, then see the label that corresponds with the highest value
factorValues(grain, modal(values(grain)))
```

16. Plot the histogram and the boxplot of the data(dem, package = "RQGIS") raster.
```{r}
# create histogram
hist(dem)

# create boxplot 
boxplot(dem)
```